\mytitle

### 2023-06-09 EXappendix.tex
Working around bugs in publishers, sigh

% Please do not edit this file, expanded-appendix.tex!
% It was generated automatically from its master file appendix.tex, by programs/expand.js
% Get everything from https://github.com/haroldthimbleby/improving-science
% 
% This file expands the LaTeX file (appendix.tex), which has complex dependencies,
%   to make this simple LaTeX file (expanded-appendix.tex)
%     that depends on nothing, apart from its standard document class.
%       This expanded file depends on no other files, as instead
%         each required file's text was input and expanded recursively - search for '% expand' in the expanded file below.
% 
% (The original file is dependent on BibTeX (.bst and .bbl files), a macro file, table of contents,
%   its own and other files' .aux files,
%     as well as various data files in the directory generated
%       etc. All these file dependencies have been removed.)
%

@@

% code to make digits into blocks of 3 (as in 1000 -> 1 000)
\def\xBlockSep{\textcolor{red}{\hskip 0.3em :}\hskip 0.3em} % International Standards are to use a small space. UK typically uses a comma
% NOTE currently, the few decimal numbers (eg 0.000 5) use hskip explicitly, since defining \xBlockSep as a comma would definitely be wrong for decimals
\newcount\xdigitBlockn
\newcount\xdigitBlocktmp
\def\xdigitBlock#1{%
        \xdigitBlockn=#1\relax
        \ifnum \xdigitBlockn < 0
                $-$%
                \xdigitBlockn = -\xdigitBlockn
        \fi
        \let \xdigitBlockpptemp \empty
        \loop \ifnum \xdigitBlockn > 999
                \xdigitBlocktmp = \xdigitBlockn
                \divide \xdigitBlockn by 1000
                \xdigitBlocktmp = \numexpr \xdigitBlocktmp - 1000*\xdigitBlockn \relax
                \edef \xdigitBlockpptemp{\xBlockSep\ifnum \xdigitBlocktmp < 100  0\ifnum \xdigitBlocktmp < 10 0\fi\fi \the\xdigitBlocktmp \xdigitBlockpptemp}%
        \repeat
        \the\xdigitBlockn
        \xdigitBlockpptemp
}

@@

\@writefile{toc}{\contentsline {section}{\numberline {7}Conclusions}{28}{section.7}\protected@file@percent }

@@

@@

This is clearly all source code, but the number \the\magicNumber\ above is likely to be some sort of relevant data, though it might be a physical constant whose value does not depend at all on \emph{this\/} experiment. The next hard-coded value mentioned in the calculation is difficult to categorize: is the value of $\pi$ empirical data or is it part of a standard formula? Some programming languages like \emph{Mathematica\/} treat $\pi$ as an exact mathematical constant (e.g., \emph{Mathematica\/} calculates $\tan\pi/4 = 1$ exactly), but $\pi$ is \emph{also\/} definitely an inexact empirical value.\footnote{A record set on 19 August 2021, the most accurate value of $\pi$ then known was 62,831,853,071,796 digits %as calculated by the University of Applied Sciences, Chur, Switzerland 
\url{www.fhgr.ch/en/specialist-areas/applied-future-technologies/davis-centre/pi-challenge}}

@@

many programmers would optimize this to \texttt{A}, because they know the condition is true because of their assumptions. This now seems to be a more efficient program because it has avoided a test (which a modern complier would have optimized away anyway). Unfortunately, the previously explicit dependency of the code on the magic number \the\magicNumber\ has completely disappeared.

@@

\begin{figure} 
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{expanded-basic.\imageFileSuffix} 
   \caption{The common basic --- error-prone and obsolete --- approach to scientific authoring is to use code to help generate analyses and diagrams, then manually copy and paste the selected results into the publication. Note that the publication, including the results, can be edited arbitrarily, and typically the results published will have been edited and modified (if only for typographical purposes) from those actually generated by the code.}
   \label{supplement-fig:basic}
\end{figure}
\begin{figure} 
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{expanded-excel.\imageFileSuffix} 
   \caption{When the code is in a spreadsheet, such as Microsoft Excel, the code is generally hidden from sight. The data copied \& pasted into a publication may or may not be calculated from data in other cells (such as column totals). Records are rarely taken of these manual processes, and, anyway, typically it is impossible to be certain exactly what has been copied unless very great care is taken. In consequence, if a spreadsheet is modified, it is haphazard what results are updated and corrected in the publication. }
   \label{supplement-fig:excel}
\end{figure}
\begin{figure} 
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{expanded-import.\imageFileSuffix} 
   \caption{Improving over the normal approach (figure \ref{supplement-fig:basic}), data and diagrams are placed in the publication using a programmed, systematic approach. In the present paper, this was done by generating text files of \LaTeX\ definitions (represented by boxes in the central column of the schematic), hence providing \LaTeX\ names for all of the code-generated values. Since the copying of results into a publication is automated and easy, any improvements to the results (such as correcting errors) are made by improving the code --- which therefore contributes to improving all future-generated results too. Not shown, but in well-engineered code, documentation will also be generated, such as by using tools such as JavaDoc or Doxygen.}
   \label{supplement-fig:rap}
\end{figure}
\begin{figure}[t]
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{expanded-notebook.\imageFileSuffix} 
   \caption{In a notebook system, such as \emph{Mathematica\/} or Jupyter, a single document --- the notebook --- integrates the publication, the code, any notes, as well as all the results. The code generates results and images that are automatically (and reliably) inserted in place into the notebook, usually right after the code that generates them. Some parts of the notebook are marked or selected to be printed as the publication, thus allowing most if not all code to be hidden from the final publication. In well-engineered code, the notebook will directly contain full documentation.
Note that in most systems, notebooks can import and generate arbitrary data (images, sounds, sensor data, etc).}
   \label{supplement-fig:notebook}
\end{figure}

@@

This convenience sample may be considered to be small given the importance of the research questions and relative to the diversity and huge number of scientific papers,\footnote{Using Google Scholar it is estimated that over 40,000 papers meeting the title criteria were published in the month of July 2020.} but~\ldots

@@

All the discussion and examples above were generated automatically, and have been checked correct for other correct values than \the\countUsesVersionControlRepository. This approach, too, considerably helps defend against common Human Factors errors. For example, if we set \texttt{\bslash countUsesVersionControlRepository}=\the\countUsesVersionControlRepository\ to be 2\hskip .3em 348, say, then all of the subsequent sentences that mention it will say something unexpected and so have to be more carefully proof-read, significantly reducing confirmation bias. The approach turns a possibly-hard-to-spot \emph{single\/} error into \emph{multiple\/} errors spread throughout the paper into different contexts, thus increasing the chances of noticing the error.

>>>>

% Please do not edit this file, EXappendix.tex!
% It was generated automatically from its master file appendix.tex, by programs/expand.js
% Get everything from https://github.com/haroldthimbleby/improving-science
% 
% This file expands the LaTeX file (appendix.tex), which has complex dependencies,
%   to make this simple LaTeX file (EXappendix.tex)
%     that depends on nothing, apart from its standard document class.
%       This expanded file depends on no other files, as instead
%         each required file's text was input and expanded recursively - search for '% expand' in the expanded file below.
% 
% (The original file is dependent on BibTeX (.bst and .bbl files), a macro file, table of contents,
%   its own and other files' .aux files,
%     as well as various data files in the directory generated
%       etc. All these file dependencies have been removed.)
%

@@

% code to make digits into blocks of 3 (as in 1000 -> 1 000)
\def\xBlockSep{\hskip 0.3em} % International Standards are to use a small space. UK typically uses a comma
% NOTE currently, the few decimal numbers (eg 0.000 5) use hskip explicitly, since defining \xBlockSep as a comma would definitely be wrong for decimals
\newcount\xdigitBlockn
\newcount\xdigitBlocktmp
\newcount\xdigitBlocktmq

\def\xdigitBlock#1{%
        \xdigitBlockn = #1 \relax
        \ifnum \xdigitBlockn < 0
                $-$%
                \xdigitBlockn = -\xdigitBlockn
        \fi
        \let \xdigitblockinternalmacro \empty
        \loop 
           \ifnum \xdigitBlockn > 999
                \xdigitBlocktmp = \xdigitBlockn
                \divide \xdigitBlockn by 1000

                % following line rewritten explicitly for old TeX 
                %\xdigitBlocktmp = \numexpr \xdigitBlocktmp - 1000*\xdigitBlockn \relax           
                \xdigitBlocktmq = \xdigitBlockn \relax
                \multiply \xdigitBlocktmq by 1000
                \advance \xdigitBlocktmp by -\xdigitBlocktmq

                \edef \xdigitblockinternalmacro{%
                	\xBlockSep
					\ifnum \xdigitBlocktmp < 100 0\ifnum \xdigitBlocktmp < 10 0\fi \fi 
					\the\xdigitBlocktmp
					\xdigitblockinternalmacro
				}%
        \repeat 
        \the\xdigitBlockn
        \xdigitblockinternalmacro
}

@@

@@

\@writefile{toc}{\contentsline {section}{\numberline {7}Conclusions}{29}{section.7}\protected@file@percent }

@@

This is clearly all source code, but the number \the\magicNumber\ above is likely to be some sort of relevant data, though it might be a physical constant whose value does not depend at all on \emph{this\/} experiment. The next hard-coded value mentioned in the calculation is difficult to categorize: is the value of $\pi$ empirical data or is it part of a standard formula? Some programming languages like \emph{Mathematica\/} treat $\pi$ as an exact mathematical constant (e.g., \emph{Mathematica\/} calculates $\tan\pi/4 = 1$ exactly), but $\pi$ is \emph{also\/} definitely an inexact empirical value.\footnote{A record set on 19 August 2021, the most accurate value of $\pi$ then known was 
62\xBlockSep 831\xBlockSep 853\xBlockSep 071\xBlockSep 796 digits. %as calculated by the University of Applied Sciences, Chur, Switzerland 
See \url{www.fhgr.ch/en/specialist-areas/applied-future-technologies/davis-centre/pi-challenge}}

@@

many programmers would optimize this to \texttt{A}, because they know the condition is true because of their assumptions. This now seems to be a more efficient program because it has avoided a test (which a modern complier would have optimized away anyway). Unfortunately, the previously explicit dependency of the code on specific the magic number \the\magicNumber\ has completely disappeared.

@@

\begin{figure} 
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{EXbasic} 
   \caption{The common basic --- error-prone and obsolete --- approach to scientific authoring is to use code to help generate analyses and diagrams, then manually copy and paste the selected results into the publication. Note that the publication, including the results, can be edited arbitrarily, and typically the results published will have been edited and modified (if only for typographical purposes) from those actually generated by the code.}
   \label{supplement-fig:basic}
\end{figure}
\begin{figure} 
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{EXexcel} 
   \caption{When the code is in a spreadsheet, such as Microsoft Excel, the code is generally hidden from sight. The data copied \& pasted into a publication may or may not be calculated from data in other cells (such as column totals). Records are rarely taken of these manual processes, and, anyway, typically it is impossible to be certain exactly what has been copied unless very great care is taken. In consequence, if a spreadsheet is modified, it is haphazard what results are updated and corrected in the publication. }
   \label{supplement-fig:excel}
\end{figure}
\begin{figure} 
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{EXimport} 
   \caption{Improving over the normal approach (figure \ref{supplement-fig:basic}), data and diagrams are placed in the publication using a programmed, systematic approach. In the present paper, this was done by generating text files of \LaTeX\ definitions (represented by boxes in the central column of the schematic), hence providing \LaTeX\ names for all of the code-generated values. Since the copying of results into a publication is automated and easy, any improvements to the results (such as correcting errors) are made by improving the code --- which therefore contributes to improving all future-generated results too. Not shown, but in well-engineered code, documentation will also be generated, such as by using tools such as JavaDoc or Doxygen.}
   \label{supplement-fig:rap}
\end{figure}
\begin{figure}[t]
   \centering
   % removed generated/ from graphics pathname
\includegraphics[width=5in]{EXnotebook} 
   \caption{In a notebook system, such as \emph{Mathematica\/} or Jupyter, a single document --- the notebook --- integrates the publication, the code, any notes, as well as all the results. The code generates results and images that are automatically (and reliably) inserted in place into the notebook, usually right after the code that generates them. Some parts of the notebook are marked or selected to be printed as the publication, thus allowing most if not all code to be hidden from the final publication. In well-engineered code, the notebook will directly contain full documentation.
Note that in most systems, notebooks can import and generate arbitrary data (images, sounds, sensor data, etc).}
   \label{supplement-fig:notebook}
\end{figure}

@@

This convenience sample may be considered to be small given the importance of the research questions and relative to the diversity and huge number of scientific papers,\footnote{Using Google Scholar it is estimated that over \digitBlock{40000} papers meeting the title criteria were published in the month of July 2020.} but~\ldots

@@

All the discussion and examples above were generated automatically, and have been checked correct for other correct values than \the\countUsesVersionControlRepository. This approach, too, helps defend against common Human Factors errors. For example, if we set \texttt{\bslash countUsesVersionControlRepository}=\the\countUsesVersionControlRepository\ to be \digitBlock{2348} instead, say, then all of the subsequent sentences that mention it will say something unexpected and so have to be more carefully proof-read, significantly reducing confirmation bias. The approach turns a possibly-hard-to-spot \emph{single\/} error into \emph{multiple\/} errors spread throughout the paper into different contexts, thus increasing the chances of noticing the error.

### 2023-06-09 EXpaper.tex
Working around bugs in publishers, sigh

% Please do not edit this file, expanded-paper.tex!
% It was generated automatically from its master file paper.tex, by programs/expand.js
% Get everything from https://github.com/haroldthimbleby/improving-science
% 
% This file expands the LaTeX file (paper.tex), which has complex dependencies,
%   to make this simple LaTeX file (expanded-paper.tex)
%     that depends on nothing, apart from its standard document class.
%       This expanded file depends on no other files, as instead
%         each required file's text was input and expanded recursively - search for '% expand' in the expanded file below.
% 
% (The original file is dependent on BibTeX (.bst and .bbl files), a macro file, table of contents,
%   its own and other files' .aux files,
%     as well as various data files in the directory generated
%       etc. All these file dependencies have been removed.)
%

@@

\received{30 April 2022}
\revised{8 June 2022, 3 January 2023, 31 January 2023, 12~May~2023 \& 7~June~2023} 
\volumenumber{00} \issuenumber{0} \volumeyear{2023}

@@

\begin{table}
\vskip 1.2ex % align top of columns!
% expand table of contents paper.toc
\makeatletter
  \section*{\contentsname
    \@mkboth{\MakeUppercase\contentsname}{\MakeUppercase\contentsname}}
\makeatother\contentsline {section}{\numberline {1}Introduction}{1}{section.1}%
\contentsline {section}{\numberline {2}Background}{2}{section.2}%
\contentsline {subsection}{\numberline {2.a}Code quality concerns}{4}{subsection.2.1}%
\contentsline {subsection}{\numberline {2.b}Computable papers}{5}{subsection.2.2}%
\contentsline {subsection}{\numberline {2.c}RAP: Reproducible Analytical Pipelines}{7}{subsection.2.3}%
\contentsline {subsubsection}{\numberline {2.c.1}Potential code}{9}{subsubsection.2.3.1}%
\contentsline {subsubsection}{\numberline {2.c.2}RAP as research}{9}{subsubsection.2.3.2}%
\contentsline {section}{\numberline {3}The statistics/code analogy}{10}{section.3}%
\contentsline {subsection}{\numberline {3.a}The siren call of over-fitting code}{11}{subsection.3.1}%
\contentsline {section}{\numberline {4}The conventional role of code}{13}{section.4}%
\contentsline {subsection}{\numberline {4.a}The deceptive simplicity of code}{14}{subsection.4.1}%
\contentsline {subsection}{\numberline {4.b}The low status of coding}{15}{subsection.4.2}%
\contentsline {subsection}{\numberline {4.c}The critical role of code is often ignored}{15}{subsection.4.3}%
\contentsline {subsection}{\numberline {4.d}Bugs, code and programming}{15}{subsection.4.4}%
\contentsline {subsection}{\numberline {4.e}Long-term problems of unreliable code}{16}{subsection.4.5}%
\contentsline {section}{\numberline {5}State of the art}{17}{section.5}%
\contentsline {subsection}{\numberline {5.a}Case study: Pandemic modeling}{17}{subsection.5.1}%
\contentsline {subsection}{\numberline {5.b}Concerns with reproducibility}{18}{subsection.5.2}%
\contentsline {subsection}{\numberline {5.c}Beyond pandemic modeling}{20}{subsection.5.3}%
\contentsline {subsection}{\numberline {5.d}The lop-sided emphasis on data}{21}{subsection.5.4}%
\contentsline {section}{\numberline {6}Rethinking science that uses code}{22}{section.6}%
\contentsline {subsection}{\numberline {6.a}Software Engineering Boards}{22}{subsection.6.1}%
\contentsline {subsection}{\numberline {6.b}Extending RAP to RAP{\tt +}}{23}{subsection.6.2}%
\contentsline {subsection}{\numberline {6.c}The paper as a scientific laboratory}{24}{subsection.6.3}%
\contentsline {subsection}{\numberline {6.d}Action must be interdisciplinary}{25}{subsection.6.4}%
\contentsline {subsection}{\numberline {6.e}Methodological statements}{25}{subsection.6.5}%
\contentsline {subsection}{\numberline {6.f}Training to reduce technical debt}{26}{subsection.6.6}%
\contentsline {subsection}{\numberline {6.g}Benefits beyond science}{27}{subsection.6.7}%
\contentsline {subsection}{\numberline {6.h}Approaches to further work}{28}{subsection.6.8}%
\contentsline {section}{\numberline {7}Conclusions}{28}{section.7}%

@@

\begin{table*}
{\sf
\newcount \enum
\enum=1
\renewcommand{\theenumi}{%\ifnum \enum<10 \hphantom{0}\fi
\the\enum
\global\advance \enum by 1}
\begin{center}\begin{tabular}{|rp{6.5in}|}
\hline 1&
Peer-review is used to ensure the workflow followed is reproducible and to identify improvements 
\\\hline 2&No or minimal manual interference;
for example copy-paste, point-click and drag-drop steps replaced using computer code that can be inspected by others 
\\\hline 3&Open-source programming languages so that processes do not rely on proprietary software and can be reproduced by others
\\\hline 4&Version control software, such as Git, to guarantee an audit trail of changes made to code 
\\\hline 5&Publication of code, whenever possible, on code hosting platforms such as GitHub to improve transparency 
\\\hline 6&Well-commented code and embedded documentation to ensure the workflow can be understood and used by others 
\\\hline 7&Embedding of existing quality assurance practices in code, following guidance set by recognized organizations 
\\\hline 
\multicolumn{2}{@{}c@{}}{\begin{minipage}{6.8in}\vskip 1ex
Adopting RAP principles is not necessarily about incorporating all of the above: implementing just some of these principles will generate valuable improvements.\end{minipage}}
\end{tabular}\end{center}\vskip -1ex
}\caption{A minimum standard of RAP, based on the UK Statistics Authority summary \cite{RAPstandard}.}
\label{figure-RAP-summary}
\end{table*}

@@

\item Another case of potential code arises when a paper has been completed, but referees or the publishers require some minor fixes. These minor fixes are easier to implement in the paper as simple textual corrections, rather than revisiting and updating the explicit code that informed or generated the paper (and updating the repositories and so forth). The RAP approach would require the entire workflow to be revised, not just the final presentation in the paper.

\end{enumerate}

@@

\begin{figure\figureStarDetails}[t]
{ \newdimen \imageWidth 
  \imageWidth=\textwidth
  \imageWidthCalculation
  \begin{center}%
% removed generated/ from graphics pathname
\includegraphics[width=\imageWidth]{expanded-plot.\imageFileSuffix}
\end{center}}
\caption{Much computational science is concerned with finding plausible multi-dimensional models that fit models to data with the aim of extrapolating or predicting new results from them. Shown here is notional sample of experimental 2D data (the dots), a linear least squares regression, and an exact polynomial model. The over-fitted polynomial model fits the sample \emph{exactly\/}, but since the experimental data is presumably subject to random error (indicated by the confidence interval, itself estimated) the linear model would generally be considered a better description of the experimental data.}
\label{fig-overfit}
\end{figure\figureStarDetails}

@@

\newcount \enum
\enum=1
\renewcommand{\theenumi}{\ifnum \enum<10 \hphantom{0}\fi
\the\enum
\global\advance \enum by 1}
\begin{enumerate}\raggedright
\item Is the code valid: does it do what the paper claims?\footnote{Of course, the underlying science may be wrong to, so it is useful to distinguish \emph{internal validity\/} and \emph{external validity}. Internal validity occurs if the code does what the paper claims; external validity occurs if the code represents correct science which the paper may have interpreted incorrectly.}
\item Do other scientists, including reviewers and the authors, understand the code?
\item Does the code implement the methods described in the paper?
\item Has the code been over-fitted or tweaked to support specific claims in the paper?
\item Is there a definitive version of code?
\item Is the code controlled and signed?
\item What limitations does the code have?
\item Was the code developed to any standard, and does it comply to that standard?
\item How does the code protect against data, coding, and human error?
\item Was the code tested adequately?
\item Does the code depend on arbitrary parameters, data, or code to over-fit to obtain the published results?
\item Is the code documented adequately, so we know what it is trying to do, and how?
\item \ldots\ and so forth. 
\end{enumerate}

@@

Finally, as there is no practical distinction between data and code (see \supplement) and methodology (thanks to RAP), and certainly no distinctions that cannot be circumvented, journal and funder policies of code and data access should be reviewed \emph{and unified\/} so that the access and methodology statements apply to all information, regardless of arbitrary classification of it as code or data (or documentation, assurance case, etc).

@@

\enum=1
\begin{enumerate}\raggedright
\itemsep=1.5ex

@@

\item
To find effective ways to promote professional software engineers being recognized and participating fully in scientific research, just as professional statisticians routinely support quality research (see section \ref{statistics-computation-analogy}).
\end{enumerate}

@@

\bibnewcolumn = 126 % bibentry number of last entry on (final) left hand column
% Hint: first set it to a very large number, so the final column is as unbalanced as it gets
% then, if it's worth fixing the final page, lower the number to some citation on the 
% lower part of the left hand column; this citation will then become the last in that
% column. Hopefully the righthand column will be of equal length!
% Set to -1 (not 0) if you don't want this feature
% PS Don't forget to check there is nothing on the following page!

>>>>

% Please do not edit this file, EXpaper.tex!
% It was generated automatically from its master file paper.tex, by programs/expand.js
% Get everything from https://github.com/haroldthimbleby/improving-science
% 
% This file expands the LaTeX file (paper.tex), which has complex dependencies,
%   to make this simple LaTeX file (EXpaper.tex)
%     that depends on nothing, apart from its standard document class.
%       This expanded file depends on no other files, as instead
%         each required file's text was input and expanded recursively - search for '% expand' in the expanded file below.
% 
% (The original file is dependent on BibTeX (.bst and .bbl files), a macro file, table of contents,
%   its own and other files' .aux files,
%     as well as various data files in the directory generated
%       etc. All these file dependencies have been removed.)
%

@@

\received{30 April 2022}
\revised{8 June 2022, 3 January 2023, 31 January 2023, 12~May~2023 \& 7~June~2023} 
\accepted{30 May 2023}
\volumenumber{00} \issuenumber{0} \volumeyear{2023}

@@

\begin{table}
\vskip 1.2ex % align top of columns!
% expand table of contents paper.toc
\makeatletter
  \section*{\contentsname
    \@mkboth{\MakeUppercase\contentsname}{\MakeUppercase\contentsname}}
\makeatother\contentsline {section}{\numberline {1}Introduction}{1}{section.1}%
\contentsline {section}{\numberline {2}Background}{2}{section.2}%
\contentsline {subsection}{\numberline {2.a}Code quality concerns}{4}{subsection.2.1}%
\contentsline {subsection}{\numberline {2.b}Computable papers}{5}{subsection.2.2}%
\contentsline {subsection}{\numberline {2.c}RAP: Reproducible Analytical Pipelines}{7}{subsection.2.3}%
\contentsline {subsubsection}{\numberline {2.c.1}Potential code}{9}{subsubsection.2.3.1}%
\contentsline {subsubsection}{\numberline {2.c.2}RAP as research}{9}{subsubsection.2.3.2}%
\contentsline {section}{\numberline {3}The statistics/code analogy}{10}{section.3}%
\contentsline {subsection}{\numberline {3.a}The siren call of over-fitting code}{11}{subsection.3.1}%
\contentsline {section}{\numberline {4}The conventional role of code}{13}{section.4}%
\contentsline {subsection}{\numberline {4.a}The deceptive simplicity of code}{14}{subsection.4.1}%
\contentsline {subsection}{\numberline {4.b}The low status of coding}{15}{subsection.4.2}%
\contentsline {subsection}{\numberline {4.c}The critical role of code is often ignored}{15}{subsection.4.3}%
\contentsline {subsection}{\numberline {4.d}Bugs, code and programming}{15}{subsection.4.4}%
\contentsline {subsection}{\numberline {4.e}Long-term problems of unreliable code}{16}{subsection.4.5}%
\contentsline {section}{\numberline {5}State of the art}{17}{section.5}%
\contentsline {subsection}{\numberline {5.a}Case study: Pandemic modeling}{17}{subsection.5.1}%
\contentsline {subsection}{\numberline {5.b}Concerns with reproducibility}{18}{subsection.5.2}%
\contentsline {subsection}{\numberline {5.c}Beyond pandemic modeling}{20}{subsection.5.3}%
\contentsline {subsection}{\numberline {5.d}The lop-sided emphasis on data}{21}{subsection.5.4}%
\contentsline {section}{\numberline {6}Rethinking science that uses code}{22}{section.6}%
\contentsline {subsection}{\numberline {6.a}Software Engineering Boards}{22}{subsection.6.1}%
\contentsline {subsection}{\numberline {6.b}Extending RAP to RAP{\tt +}}{23}{subsection.6.2}%
\contentsline {subsection}{\numberline {6.c}The paper as a scientific laboratory}{24}{subsection.6.3}%
\contentsline {subsection}{\numberline {6.d}Action must be interdisciplinary}{25}{subsection.6.4}%
\contentsline {subsection}{\numberline {6.e}Methodological statements}{25}{subsection.6.5}%
\contentsline {subsection}{\numberline {6.f}Training to reduce technical debt}{26}{subsection.6.6}%
\contentsline {subsection}{\numberline {6.g}Benefits beyond science}{27}{subsection.6.7}%
\contentsline {subsection}{\numberline {6.h}Approaches to further work}{28}{subsection.6.8}%
\contentsline {section}{\numberline {7}Conclusions}{29}{section.7}%

@@

\begin{table*}
{\sf
\begin{center}\begin{tabular}{|rp{6.5in}|}
\hline 1&
Peer-review is used to ensure the workflow followed is reproducible and to identify improvements 
\\\hline 2&No or minimal manual interference;
for example copy-paste, point-click and drag-drop steps replaced using computer code that can be inspected by others 
\\\hline 3&Open-source programming languages so that processes do not rely on proprietary software and can be reproduced by others
\\\hline 4&Version control software, such as Git, to guarantee an audit trail of changes made to code 
\\\hline 5&Publication of code, whenever possible, on code hosting platforms such as GitHub to improve transparency 
\\\hline 6&Well-commented code and embedded documentation to ensure the workflow can be understood and used by others 
\\\hline 7&Embedding of existing quality assurance practices in code, following guidance set by recognized organizations 
\\\hline 
\multicolumn{2}{@{}c@{}}{\begin{minipage}{6.8in}\vskip 1ex
Adopting RAP principles is not necessarily about incorporating all of the above: implementing just some of these principles will generate valuable improvements.\end{minipage}}
\end{tabular}\end{center}\vskip -1ex
}\caption{A minimum standard of RAP, based on the UK Statistics Authority summary \cite{RAPstandard}.}
\label{figure-RAP-summary}
\end{table*}

@@

\item Another case of potential code arises when a paper has been completed, but referees or the publishers require some minor fixes. These minor fixes are easier to implement in the paper as simple textual corrections, rather than revisiting and updating the explicit code that informed or generated the paper (and updating the repositories and so forth). The RAP approach would require the entire workflow to be revised, not just the final presentation in the paper.
\end{enumerate}

@@

\begin{figure\figureStarDetails}[t]
{ \newdimen \imageWidth 
  \imageWidth=\textwidth
  \imageWidthCalculation
  \begin{center}%
% removed generated/ from graphics pathname
\includegraphics[width=\imageWidth]{EXplot}
\end{center}}
\caption{Much computational science is concerned with finding plausible multi-dimensional models that fit models to data with the aim of extrapolating or predicting new results from them. Shown here is notional sample of experimental 2D data (the dots), a linear least squares regression, and an exact polynomial model. The over-fitted polynomial model fits the sample \emph{exactly\/}, but since the experimental data is presumably subject to random error (indicated by the confidence interval, itself estimated) the linear model would generally be considered a better description of the experimental data.}
\label{fig-overfit}
\end{figure\figureStarDetails}

@@

\begin{enumerate}
\raggedright
\item Is the code valid: does it do what the paper claims?\footnote{Of course, the underlying science may be wrong to, so it is useful to distinguish \emph{internal validity\/} and \emph{external validity}. Internal validity occurs if the code does what the paper claims; external validity occurs if the code represents correct science which the paper may have interpreted incorrectly.}
\item Do other scientists, including reviewers and the authors, understand the code?
\item Does the code implement the methods described in the paper?
\item Has the code been over-fitted or tweaked to support specific claims in the paper?
\item Is there a definitive version of code?
\item Is the code controlled and signed?
\item What limitations does the code have?
\item Was the code developed to any standard, and does it comply to that standard?
\item How does the code protect against data, coding, and human error?
\item Was the code tested adequately?
\item Does the code depend on arbitrary parameters, data, or code to over-fit to obtain the published results?
\item Is the code documented adequately, so we know what it is trying to do, and how?
\item \ldots\ and so forth. 
\end{enumerate}

@@

As there is no practical distinction between data and code (see \supplement) and methodology (thanks to RAP), and certainly no distinctions that cannot be circumvented, journal and funder policies of code and data access should be reviewed \emph{and unified\/} so that the access and methodology statements apply to all information, regardless of arbitrary classification of it as code or data (or documentation, assurance case, etc).

Finally, there is still a long way to go. The \emph{Computer Journal}, which should be a leading journal for publishing computational science (and which has published this paper) uses poorly documented automated and buggy paper processing software, whose obscure bugs took several weeks of hard work to overcome in order to submit the present paper. Such problems will put authors off using computational papers and \RAPstar\ methodologies.

@@

\begin{enumerate}\raggedright
\itemsep=1.5ex

@@

\item
To find effective ways to promote professional software engineers being recognized and participating fully in scientific research, just as professional statisticians routinely support quality research (see section \ref{statistics-computation-analogy}).

\item
Computer science is a complex discipline, but with low barriers to entry. When critical code is being written or used, high quality developers and software engineers are required, and there are not many of them. Improving the educational and professional standards in coding, and awareness of the issues are a priority for all aspects of science, publishing --- and life generally.
\end{enumerate}

@@

\bibnewcolumn = 123 % bibentry number of last entry on (final) left hand column
% Hint: first set it to a very large number, so the final column is as unbalanced as it gets
% then, if it's worth fixing the final page, lower the number to some citation on the 
% lower part of the left hand column; this citation will then become the last in that
% column. Hopefully the righthand column will be of equal length!
% Set to -1 (not 0) if you don't want this feature
% PS Don't forget to check there is nothing on the following page!
